---
title: "Bayesian LCA"
output: html_document
date: '2022-07-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Latent Class Analysis

This latent class analysis project, we assume the indicator variables are dichotomous. 

### Simulating the dichotomous variables
Suppose there are $M$ dichotomously scored items $(Y_1, Y_2, Y_3, \ldots, Y_M)$ observed on $N$ participants for measuring a latent trait, $\theta$. Based on the responses to these categorical indicators, the participants can be categorized into $c$ classes. 
Suppose that the $c = \{2, 3, \dots, K\}$ class have marginal probabilities $P = \{p_1, p_2, \ldots, p_K \}$ such that $\sum_{c = 1}^{K} p_c = 1$

```{r}
library(data.table)
LCA_sim <- function(NC, NI, N, p, cp){
  p <- p/sum(p)
  cl <- sample(x = 1:NC, size = N, replace = TRUE, prob = p)
  y <- matrix(nrow = N, ncol = NI)
  
  for(i in 1:N){
    for(j in 1:NC) {
      if(cl[i] == j){
        y[i, ] <- rbinom(n = rep(1, NI), size = rep(1, NI), prob = cp[j,])
      }
    }
  }
  Y <- as.data.frame(y)
  colnames(Y)<- paste0("V", 1:NI)
  Y$group <- cl
  return(Y)
}

p <- c(.21, .62, .17)

cp <- matrix(data = c(.7, .35, .05, .09, .31, .06, .01, .05, .72, .49, .3, .62), 
             ncol = 4, nrow = 3, byrow = TRUE)

cp
```

Simulating a dataset 
```{r}
Y <- LCA_sim(NC = 3, NI = 4, N = 500, p = p, cp = cp)
Y <- data.table(Y)

Y[, lapply(.SD, mean), by = group, .SDcols = paste0("V", 1:4)][order(group),]
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
